---
format: gfm
---

```{r setup, echo=FALSE}

suppressPackageStartupMessages(library(tidyverse))

library(bigrquery)

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE, 
  dpi = 600
)
options(scipen = 999, digits = 2)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r connection}
#| echo: false
#' BigQuery 
bq_con <- dbConnect(
  bigrquery::bigquery(),
  project = "subugoe-collaborative",
  billing = "subugoe-collaborative"
)

jct_esac <- tbl(bq_con, "openbib.jct_esac")
jct_journals <- tbl(bq_con, "openbib.jct_journals")
jct_institutions <- tbl(bq_con, "openbib.jct_institutions")
jct_articles <- tbl(bq_con, "openbib.jct_articles")
```

# Transformative Agreements Dataset

## Overview

This dataset preserves historical records of the  [public Transformative Agreement Data dumps](https://journalcheckertool.org/transformative-agreements/) as provided by the [Journal Checker Tool](https://journalcheckertool.org/) from the [cOAlition S](https://www.coalition-s.org/) since July 2021.
By linking agreement data with OpenAlex and using standardised identifiers (ISSN-L for journals, ROR-IDs for institutions), it enables open data analyses on these agreements' evolution and impact over time.

Overall, this dataset comprises:

- `r jct_esac |> filter(jct_jn == TRUE) |> distinct(id) |> collect() |> nrow()` transformative agreements with data about `r jct_journals |> distinct(issn_l) |> collect() |> nrow()` eligible journals and `r jct_institutions |> distinct(ror_id) |> collect() |> nrow()` participating institutions. 
- `r jct_articles |> distinct(doi) |> collect() |> nrow()` eligible articles published by first authors affiliated with participating institutions.

## Motivation 

Quantitative science studies are increasingly examining the impact of transformative agreements on scholarly publishing. As transformative agreements have also become a much debated licensing model for libraries, measuring their extent is also high on the policy agenda. Unfortunately, information on this licensing model is only partially available and scattered across different open data sources.

For example, the [public Transformative Agreement Data dumps](https://journalcheckertool.org/transformative-agreements/) as provided by the [Journal Checker Tool](https://journalcheckertool.org/) from the [cOAlition S](https://www.coalition-s.org/) (JCT), which is the most comprehensive source for journals and institutions covered by transformative agreements, provides data only for current agreements; information about expired agreements is not available, limiting longitudinal studies.

To address this data gap, this dataset combines open data from different sources to improve the transparency of transformative agreements and to allow estimates about articles published under these agreements. It uses JCT transformative agreement data for information on journal portfolios and participating institutions. The data were normalised using ISSN-L for journals (as provided by the ISSN Agency) and associated ROR-IDs (as provided by OpenAlex). Due to limited public invoicing data, articles published under transformative agreements were identified by linking first author affiliations from OpenAlex to eligible institutions according to agreement data.

As such, the dataset aims to support data-driven research and monitoring of open access business models and transformative agreements using open data and methods.

## Methods

### Data retrieval

A dedicated [bot](https://github.com/njahn82/jct_data/) has been fetching the cOAlition S Public Transformative Agreement Data dump serving the Journal Checker Tool weekly since December 2022. 
The resulting data snapshots were saved using Git and are available via [GitHub](https://github.com/njahn82/jct_data?tab=readme-ov-file#snapshots). 
To create a comprehensive dataset, the weekly snapshots were retrieved from git and merged together using a custom  [script](https://github.com/njahn82/jct_data/blob/main/combine.R)). As the Transformative Agreement Data are continuously curated, only the most recent data about an agreement were safeguarded. For expired agreements, this represented the latest snapshot available.

### Data curation

The cOAlition S Public Transformative Agreement Data dumps link agreements to journals represented by journal names and ISSN. After mapping ISSN variants to the corresponding linking ISSN (ISSN-L), as provided by the ISSN International Centre, journals were associated with publishers according to the [ESAC Transformative Agreement Registry](https://esac-initiative.org/about/transformative-agreements/agreement-registry/). This reflects that some portfolios may include imprints.

cOAlition S Public Transformative Agreement Data did not comprehensively cover associated institutions, such as university hospitals or institutes of large research organisations like the Max Planck Society. To improve matching, cOAlition S Public Transformative Agreement Data were automatically enriched with ROR-IDs from associated organisations according to OpenAlex's institution entity data.

To estimate articles enabled by transformative agreements, article metadata were retrieved for each journal from OpenAlex. Then, participating institutions were matched with the first author affiliations recorded by OpenAlex. The matching considered first author affiliations instead of corresponding author affiliation due to the lack of corresponding author data in OpenAlex. Matching also considered the duration of agreements according to the [ESAC registry](https://esac-initiative.org/about/transformative-agreements/agreement-registry/).

The matching was carried out on [Google BigQuery](https://console.cloud.google.com/bigquery) for performance reasons.

The compilation of the datasets was carried out in April 2025 using the most recent datasets.

## Data files

The dataset consists of four files:

**Historic cOAlition S Transformative Agreement Data** 

- `jct_journals` links journals (represented by ISSN-L) to transformative agreements (represented by ESAC ID)
- `jct_institutions` links participating institutions (represented by ROR-ID) to transformative agreements (represented by ESAC ID)

**ESAC snapshot**

- `jct_esac` - Metadata about the transformative agreements from the ESAC registry, in particular publisher name and duration of agreement

**Articles under Transformative Agreements**

- `jct_articles` links articles indexed in OpenAlex to transformative agreements through matching with `jct_journals` (by ISSN-L), `jct_institutions` (by ROR-ID representing first-author afilliation) and `jct_esac` (by ESAC ID and duration of agreement)

You can find the documentation of data files in the [data documentation](data_overview.md#-transformative-agreements)

## Use case

### How many articles were enabled by transformative agreements?

The following SQL query shows how to retrieve annual counts for articles enabled by transformative agreements. Using `jct_articles` as a starting point, retrieval is restricted to articles and reviews according to OpenAlex. Furthermore, articles were matched to publishers according to the ESAC registry.


```{sql}
#| connection: bq_con
#| output.var: bq_df
SELECT
  publication_year,
  esac.publisher,
  COUNT(DISTINCT(jct.id)) AS n,
FROM
  `subugoe-collaborative.openbib.jct_articles` AS jct
INNER JOIN
  `subugoe-collaborative.openalex.works` AS oalex
ON
  oalex.doi = jct.doi
INNER JOIN
  `subugoe-collaborative.openbib.jct_esac` AS esac
ON
  esac.id = jct.esac_id
WHERE oalex.type IN ('article', 'review') AND is_paratext = FALSE
GROUP BY
  publication_year,
  publisher
ORDER BY
  publication_year DESC,
  n DESC
```

```{r}
bq_df
```

```{r, echo =FALSE}
#| label: fig-all
#| fig-cap: "Growth of articles enabled by transformative agreements between 2020 and 2024, showing the dominance of the five largest commercial publishers in the scholarly publishing market."
pub_by_year_df <- bq_df |>
  mutate(publisher_normalised = if_else(
    publisher %in% c("Elsevier", "Springer Nature", "Wiley", "Taylor & Francis", "Sage"), publisher, "Other"
  )) |>
  group_by(publisher_normalised, publication_year) |>
  summarise(n = sum(n)) |>
  filter(between(publication_year, 2020, 2024)) |>
  mutate(publisher_normalised = factor(publisher_normalised, levels = c("Elsevier", "Springer Nature", "Wiley", "Taylor & Francis", "Sage", "Other"))) 

pub_by_year_df |>
  ggplot(aes(publication_year, n, fill = fct_rev(publisher_normalised))) +
  geom_bar(stat = "identity", color = "white", linewidth = .25) +
  scale_fill_manual("", values = c('#cccccca0', '#f0f9e8','#bae4bc','#7bccc4','#43a2ca','#0868ac')) +
  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
  labs(x = "Publication year", y = "First-author articles", title = "Growth of Transformative Agreements",
       subtitle = "Annual first-author articles from participating institutions",
       caption = "KBOPENBIB | April 2025 | CC0") +
 theme_minimal() +
  theme(plot.margin = margin(30, 30, 30, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank()) +
  theme(plot.title.position = "plot") +
  theme(plot.caption.position = "plot")

```

Figure 1 shows the growth of articles enabled by transformative agreements between 2020 and 2024. It furthermore highlights the dominance of the five large commercial publishers, with Elsevier, Springer Nature and Wiley leading.

### How many articles were made open access by transformative agreements?

Transformative agreements are complex. For instance, journal bundles can include full open access journals, hybrid journals and subscription journals. Furthermore, not all document types are allowed. Sometimes agreements also cap the number of articles that can be made open access. Additionally, authors can opt out of open access publication.

Therefore, additional open access evidence is needed in order to estimate the number of open access articles enabled by transformative agreements. The following SQL shows that article-level open access status information from OpenAlex was added.

```{sql}
#| connection: bq_con
#| output.var: bq_oa_df

SELECT
  publication_year,
  esac.publisher,
  oalex.open_access.oa_status,
  COUNT(DISTINCT(jct.id)) AS n,
FROM
  `subugoe-collaborative.openbib.jct_articles` AS jct
INNER JOIN
  `subugoe-collaborative.openalex.works` AS oalex
ON
  oalex.doi = jct.doi
INNER JOIN
  `subugoe-collaborative.openbib.jct_esac` AS esac
ON
  esac.id = jct.esac_id
WHERE oalex.type IN ('article', 'review') AND is_paratext = FALSE
GROUP BY
  publication_year,
  publisher, 
  oalex.open_access.oa_status
ORDER BY
  publication_year DESC,
  n DESC
```

```{r}
bq_oa_df
```

```{r}
#| label: fig-oa
#| echo: false
#| fig-cap: "Articles covered by transformative agreements by open access status"
oa_df <- bq_oa_df |>
  filter(oa_status %in% c("gold", "hybrid", "closed")) |>
  mutate(publisher_normalised = if_else(
    publisher %in% c("Elsevier", "Springer Nature", "Wiley", "Taylor & Francis", "Sage"), publisher, "Other"
  )) |>
  group_by(publisher_normalised, publication_year, oa_status) |>
  summarise(n = sum(n)) |>
  filter(between(publication_year, 2020, 2024)) |>
  mutate(publication_year = as.character(as.numeric(publication_year))) |>
  mutate(oa_status = factor(oa_status, levels = c("gold", "hybrid", "closed"))) |>
  mutate(publisher_normalised = factor(publisher_normalised, levels = c("Elsevier", "Springer Nature", "Wiley", "Taylor & Francis", "Sage", "Other")))
  
oa_df |>
  ggplot(aes(gsub("^20", "'", publication_year), n, fill = fct_rev(publisher_normalised))) +
  geom_bar(stat = "identity", color = "white", linewidth = .25) +
  scale_fill_manual("", values = c('#cccccca0', '#f0f9e8','#bae4bc','#7bccc4','#43a2ca','#0868ac')) +
  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
  facet_grid(~ oa_status) +
  labs(x = "Publication year", y = "First-author articles", title = "Growth of Transformative Agreements",
       subtitle = "Annual first-author articles from participating institutions, by open access status",
       caption = "KBOPENBIB | April 2025 | CC0") +
 theme_minimal() +
  theme(plot.margin = margin(30, 30, 30, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank()) +
  theme(plot.title.position = "plot") +
  theme(plot.caption.position = "plot")
```

Figure 2 shows open access by business model, highlighting that the majority of articles was made available in hybrid journals. But also the number of closed articles is striking. These can not only signal shortcomings in the matching, but also complexities of transformative agreements in general with regard to journal inclusion, article caps and the restriction of some document types. Also authors can opt out of open access.

## Responsible use

It is important to note that the  [public Transformative Agreement Data dumps](https://journalcheckertool.org/transformative-agreements/) as provided by the [Journal Checker Tool](https://journalcheckertool.org/) from the [cOAlition S](https://www.coalition-s.org/) and the ESAC Registry are voluntary, crowd-sourcing efforts. Information is therefore always subject to change.

Because of a lack of publicly available invoicing data, the provided list of articles enabled by transformative agreements represents only estimates. These estimates are based on first-author affiliations according to OpenAlex. Note that, transformative agreement guidelines require data about corresponding authors, an information, which is still not fully avaialble in OpenAlex. However, a [strong alignment between first and corresponding authorship can be observed](https://arxiv.org/abs/2504.15038v1).

## Further reading

For an coverage analysis between Scopus, Web of Science and open metadata with a specific focus on hybrid journals in transformative agreements, see:

Jahn, N. (2025). *Estimating transformative agreement impact on hybrid open access: A comparative large-scale study using Scopus, Web of Science and open metadata*. arXiv. <https://arxiv.org/abs/2504.15038>

For a validation study using internal research information, see:

de Jonge, H., Kramer, B., & Sondervan, J. (2025). *Tracking transformative agreements through open metadata: Method and validation using Dutch Research Council NWO funded papers*. MetaArXiv. <https://doi.org/10.31222/osf.io/tz6be_v1>

For an analysis of hybrid open access through transformative agreements using this data approach, see: 

Jahn, N. (2025). How open are hybrid journals included in transformative agreements? *Quantitative Science Studies*, 6, 242–262. <https://doi.org/10.1162/qss_a_00348>

For an overview about the transpareny of open access infrastructures, see:

Kramer, B. (2024). *Study on scientific publishing in Europe: Development, diversity, and transparency of costs*. Publications Office of the European Union. <https://doi.org/10.2777/89349>

For an analysis of the ESAC registry, see:

Rothfritz, L., Schmal, W. B., & Herb, U. (2024). *Trapped in transformative agreements? A multifaceted analysis of >1,000 contracts*. arXiv. https://arxiv.org/abs/2409.20224

For an overview of transformative agreement from the perspective of library consortia, see:

Dér, Á. (2025). What gets missed in the discourse on transformative agreements. *Katina Magazine*. https://doi.org/10.1146/katina-20250212-1

For economic discussion of transformative agreements, see: 

Haucap, J., Moshgbar, N., & Schmal, W. B. (2021). The impact of the German “DEAL” on competition in the academic publishing market. *Managerial and Decision Economics*, 42(8), 2027–2049. <https://doi.org/10.1002/mde.3493>

McCabe, M. J., & Mueller-Langer, F. (2024). Open access is shaping scientific communication. *Science*, 385(6714), 1170–1172. <https://doi.org/10.1126/science.adp8882>

## Contact

In case of questions and issues, please contact [Najko Jahn](mailto:najko.jahn@sub.uni-goettingen.de), SUB Göttingen.

## License

CCO

